Working title: Highly resolved spatiotemporal simulations for exploring mixed fishery dynamics

### 2019-09-16 write_up/mixfishsim_ecolmod.tex
Text revisions

\author[1]{Cóilín Minto}
\author[3]{Jean-Marc Guarini}
\author[4]{Jan Jaap Poos}

\address[1]{Galway-Mayo Institute of Technology (GMIT), Dublin Road, Galway,
	Ireland} 
\address[2]{Centre for Environment, Fisheries and Aquaculture Science (Cefas),
	Pakefield Road, Lowestoft, UK}
\address[3]{\replaced[id=JM]{Sorbonne Université, Faculty of
		Sciences}{Université Pierre et Marie Curie}, 4 Place Jussieu,
	75005 Paris, France}
\address[4]{Wageningen Marine Research, Haringkade 1 1976 CP IJmuiden,
	Netherlands}

@@

For our implementation of a random walk directional change is based on a
negatively correlated circular distribution where a favourable fishing ground
is likely to be ``fished back over" by the vessel returning in the direction it
came from.  The step length (i.e. the distance travelled from the current to
the next fishing location) is determined by relating recent
fishing success, measured as the summed value of fish caught (revenue, $Rev$);
\begin{equation}
Rev = \sum_{p=1}^{P} \replaced[id=PD]{L}{C}_{p} \cdot Pr_{p} 
\end{equation}
where $L_{p}$ is landings of a population $p$, and $Pr_{p}$ price of a
population. All species prices were kept the same across fleets and seasons.
Here, when fishing is successful vessels remain in a similar location and
continue to exploit the local fishing grounds. When unsuccessful, they move
some distance away from the current fishing location.  The movement distance
retains some degree of stochasticity, that can be controlled separately, but is
determined by the relationship: 
\begin{equation}
	StepL = \mathrm{e}^{log(\beta_{1}) + log(\beta_{2}) -
		\left(log\left(\frac{\beta_{1}}{\beta_{3}}\right)\right)} \cdot Rev
\end{equation}
Where $\beta_{1}$, $\beta_{2}$ and $\beta_{3}$ are parameters determining the
shape of the step function in its relation to revenue, so that, a step from
(x1,y1) to (x2, y2) is defined by:
\begin{equation}
	\begin{split}
 (x2, y2) =  & x1 + L \cdot \cos \left(\frac{\pi \cdot Br}{180}\right), \\
             & y1 + L \cdot \sin \left(\frac{\pi \cdot Br}{180}\right) \\	
 with  \hspace{0.5cm}     & Br_{t-1} < 180, Br_{t} = 180 + \sim vm[(0,360), k] \\
 			  & Br_{t-1} > 180, Br_{t} = 180 - \sim vm[(0,360), k] \\
	\end{split}
\end{equation}
where $L$ is the step length, $k$ the concentration parameter from the von
Mises distribution that we correlate with the revenue so
that $k = (Rev + 1 / RefRev) * max_{k}$, where $max_{k}$ is the maximum
concentration value, $k$, and $RefRev$ is parametrised as for $\beta_{3}$ in
the step length function. A realised example of the step length and turning
angle relationships to revenue can be seen at Figure \ref{S-fig:15}.

@@

We parametrised the simulation model for four example populations with
different demographics, growth rates, natural mortality and recruitment
parameters (Table \ref{tab:1}). Habitat preference (Figure \ref{S-fig:1}) and
temperature tolerances (Figures \ref{S-fig:3}, \ref{S-fig:4}) were defined to
be unique to each population resulting in differently weekly distribution
patterns (Figures \ref{S-fig:5}-\ref{S-fig:7}). In addition, each of the
populations was assumed to have two defined spawning areas that result in the
populations moving towards these areas in pre-defined weeks (Figure
\ref{S-fig:2}) with population-specific movement rates (Table \ref{tab:1}). In
such a configuration, the individual habitat preferences and thermal tolerances
result in different spatial habitat use for each population (Figure
\ref{fig:9}) and consequently different seasonal exploitation patterns (Fishing
mortality in Figure \ref{fig:10}).

@@

It can be seen from a single vessels movements during a trip that the vessel
exploits four different fishing grounds, three of them multiple times (Figure
\ref{S-fig:11}), while across several trips fishing grounds that are further
apart are fished (Figure \ref{S-fig:12}). These different locations relate to
areas where the highest revenue were experienced, as shown by
\added[id=PD]{Figure} \ref{fig:13}, where several trips for the vessel overlaid
on the revenue field, i.e. $$\sum^c_{c=1}\sum^s_{s=1} B_{s,c} \cdot Q_{s,c}$$

Vessels from the same fleet (and therefore targeting preference) may exploit
some shared and some different fishing grounds depending on their own personal
experience during the exploratory phase of the fishery (Figure \ref{S-fig:14}).
This results from the randomness in the correlated random walk step function,
with distance moved during the exploitation phase and the direction
stochastically related to the revenue experienced on the fishing ground (Figure
\ref{S-fig:15}).

@@

Figure \ref{fig:1} shows the aggregated catch composition from each of the data
sources over a ten-year period (to average seasonal patterns) at different
spatial resolutions. The finer spatial grid for the real population (top left)
and commercial data (top middle) show visually similar patterns, though there
are large unsampled areas in the commercial data from a lack of fishing
activity (particularly in the lower left part of the sampling domain). The
survey data at this spatial resolution displays very sparse information about
the spatial distributions of the populations. The slightly aggregated data on a
5 x 5 grid shows similar patterns and, while losing some of the spatial detail,
there remains good consistency between the real population and the commercial
data. Survey data starts to pick out some of the similar patterns as the other
data sources, but lacks spatiotemporal coverage. The spatial catch information
on a 10 x 10 and 20 x 20 grid lose a significant amount of information about
the spatial resolutions for all data sources, and some differences between the
survey, commercial and real population data emerge. \\

Figure \ref{fig:2} shows the consequences of different temporal aggregations of
the data over a ten-year period, with weekly (top), monthly (middle) and yearly
(bottom) catch compositions from across an aggregated 20 x 20 area. In the real
population, the monthly aggregation captures the major patterns of composition
seen in the weekly data with the percentage of different species in the catch
having similar mean and standard deviations.  In the weekly data species 1 =
9.36 (3.99), species 2 = 83.2 (5.60), species 3 = 3.57 (1.23), species 4 = 3.91
(1.59); in the monthly data species 1 = 9.23 (3.87), species 2 = 83.3 (5.52),
species 3 = 3.62 (1.15), species 4 = 3.86 (1.52). While means were similar some
of the variation was lost when aggregated to an annual level; species 1 = 9.90
(0.173), species 2 = 82.2 (0.308), species 3 = 3.82 (0.119), species 4 = 4.03
(0.0502).\\

@@

@@

\begin{figure}[!ht]
	\includegraphics[width=\linewidth]{../analysis/Data_Aggregation_space_Rev}
	\caption{Data aggregation at different spatial resolutions over a
		ten year period}
	\label{fig:1}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width =\linewidth]{../analysis/Data_Aggregation_time_Rev}
	\caption{Data aggregation at different temporal resolutions over a
		ten-year period}
	\label{fig:2}
\end{figure}

>>>>

\author[1]{Cóilín Minto}
\author[3]{Jean-Marc Guarini}
\author[4,5]{Jan Jaap Poos}

\address[1]{Galway-Mayo Institute of Technology (GMIT), Dublin Road, Galway,
	Ireland} 
\address[2]{Centre for Environment, Fisheries and Aquaculture Science (Cefas),
	Pakefield Road, Lowestoft, UK}
\address[3]{\replaced[id=JM]{Sorbonne Université, Faculty of
		Sciences}{Université Pierre et Marie Curie}, 4 Place Jussieu,
	75005 Paris, France}
\address[4]{Wageningen Marine Research, Haringkade 1 1976 CP IJmuiden,
	Netherlands}
\address[5]{Aquaculture and Fisheries Group, Wageningen University \& Research,
	Zodiac Building 122, De Elst 1, 6708 WD Wageningen, the Netherlands}

@@

For our implementation of a random walk directional change is based on a
negatively correlated circular distribution where a favourable fishing ground
is likely to be ``fished back over" by the vessel returning in the direction it
came from.  The step length (i.e. the distance travelled from the current to
the next fishing location) is determined by relating recent
fishing success, measured as the summed value of fish caught (revenue, $Rev$);
\begin{equation}
Rev = \sum_{p=1}^{P} \replaced[id=PD]{L}{C}_{p} \cdot Pr_{p} 
\end{equation}
where $L_{p}$ is landings of a population $p$, and $Pr_{p}$ price of a
population. All species prices were kept the same across fleets and seasons.
Here, when fishing is successful vessels remain in a similar location and
continue to exploit the local fishing grounds. When unsuccessful, they move
some distance away from the current fishing location.  The movement distance
retains some degree of stochasticity, that can be controlled separately, but is
determined by the relationship: 
\begin{equation}
	StepL = \mathrm{e}^{log(\beta_{1}) + log(\beta_{2}) -
		\left(log\left(\frac{\beta_{1}}{\beta_{3}}\right)\right)} \cdot Rev
\end{equation}
Where $\beta_{1}$, $\beta_{2}$ and $\beta_{3}$ are parameters determining the
shape of the step function in its relation to revenue, so that, a step from
(x1,y1) to (x2, y2) is defined by:
\begin{equation}
	\begin{split}
 (x2, y2) =  & x1 + L \cdot \cos \left(\frac{\pi \cdot Br}{180}\right), \\
             & y1 + L \cdot \sin \left(\frac{\pi \cdot Br}{180}\right) \\	
 with  \hspace{0.5cm}     & Br_{t-1} < 180, Br_{t} = 180 + \sim vm[(0,360), k] \\
 			  & Br_{t-1} > 180, Br_{t} = 180 - \sim vm[(0,360), k] \\
	\end{split}
\end{equation}
where $L$ is the step length, $k$ the concentration parameter from the von
Mises distribution that we correlate with the revenue so
that $k = (Rev + 1 / RefRev) * max_{k}$, where $max_{k}$ is the maximum
concentration value, $k$, and $RefRev$ is parametrised as for $\beta_{3}$ in
the step length function.

@@

We parametrised the simulation model for four example populations with
different demographics, growth rates, natural mortality and recruitment
parameters (Table \ref{tab:1}). Habitat preference (Figure \ref{S-fig:1}) and
temperature tolerances (Figures \ref{S-fig:3}, \ref{S-fig:4}) were defined to
be unique to each population resulting in differently weekly distribution
patterns (Figures \ref{S-fig:5}-\ref{S-fig:7}). In addition, each of the
populations was assumed to have two defined spawning areas that result in the
populations moving towards these areas in pre-defined weeks (Figure
\ref{S-fig:2}) with population-specific movement rates (Table \ref{tab:1}).

@@

Individual habitat preferences and thermal tolerances result in different
spatial habitat use for each population (Figure \ref{fig:9}) and consequently
different seasonal exploitation patterns (Fishing mortality in Figure
\ref{fig:10}).

It can be seen from a single vessels movements during a trip that the vessel
exploits three different fishing grounds, each of them multiple times (Figure
\ref{fig:mv}(A)), while across several trips fishing grounds that are further
apart are fished (Figure \ref{fig:mv} (B)). These different locations relate to
areas where the highest revenue were experienced, as shown by
\added[id=PD]{Figure} \ref{fig:13}, where several vessels tracks are
overlaid on the revenue field, i.e. $$\sum^c_{c=1}\sum^s_{s=1} B_{s,c} \cdot
Q_{s,c}$$

Vessels from the same fleet (and therefore targeting preference) may exploit
some shared and some different fishing grounds depending on their own personal
experience during the exploratory phase of the fishery (Figure \ref{fig:mv} (C)).
This results from the randomness in the correlated random walk step function,
with distance moved during the exploitation phase and the direction
stochastically related to the revenue experienced on the fishing ground (Figure
\ref{fig:mv} (D)).

@@

The aggregated catch composition from each of the data sources over a ten-year
period (which shows average seasonal patterns) at different spatial resolutions
highlights different patterns in perceived community structure depending on the
data source and aggregation level (Figure \ref{fig:1}). The finer spatial grid
for the real population (top left) and commercial data (top middle) show
visually similar patterns, though there are large unsampled areas in the
commercial data from a lack of fishing activity (particularly in the lower left
part of the sampling domain). The survey data at this spatial resolution
displays very sparse information about the spatial distributions of the
populations. The slightly aggregated data on a 5 x 5 grid shows similar
patterns and, while losing some of the spatial detail, there remains good
consistency between the real population and the commercial data. Survey data
starts to pick out some of the similar patterns as the other data sources, but
lacks spatiotemporal coverage. The spatial catch information on a 10 x 10 and
20 x 20 grid lose a significant amount of information about the spatial
resolutions for all data sources, and some differences between the survey,
commercial and real population data emerge. \\

Different perceptions of the proportion of each stock in an area are seen when
we aggregate the data over a smaller geographical region at different
timescales, with weekly (top), monthly (middle) and yearly (bottom) catch
compositions from across an aggregated 20 x 20 area (Figure \ref{fig:2}). In
the real population, the monthly aggregation captures the major patterns of
composition seen in the weekly data with the percentage of different species in
the catch having similar mean and standard deviations.  In the weekly data
species 1 = 9.36 (3.99), species 2 = 83.2 (5.60), species 3 = 3.57 (1.23),
species 4 = 3.91 (1.59); in the monthly data species 1 = 9.23 (3.87), species 2
= 83.3 (5.52), species 3 = 3.62 (1.15), species 4 = 3.86 (1.52). While means
were similar some of the variation was lost when aggregated to an annual level;
species 1 = 9.90 (0.173), species 2 = 82.2 (0.308), species 3 = 3.82 (0.119),
species 4 = 4.03 (0.0502).\\

@@

\begin{figure}[!ht]
	\includegraphics[width = \linewidth]{Plots/Combined_Movement}
	\caption{(A) The fishing locations (points) and movements (lines) of a
	single vessel during a trip overlaid on the relative value of a fishing
	site; (B) the fishing locations of the vessel over several trips (value
field changes over the period so not shown); (C) the locations of multiple
vessels from the same fleet overlaid on the value field, (D) the realised step
distance and turning angles for a single vessel over the simulation}
	\label{fig:mv}
\end{figure}

@@

\begin{figure}[!ht]
	\includegraphics[width=\linewidth]{../analysis/Data_Aggregation_space_Rev2}
	\caption{Data aggregation at different spatial resolutions over a
		ten year period}
	\label{fig:1}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width=\linewidth]{../analysis/Data_Aggregation_time_Rev2}
	\caption{Data aggregation at different temporal resolutions over a
		ten-year period}
	\label{fig:2}
\end{figure}
